---
title: "Final Shiny Project"
author: "Anand Kumar Subramaniam"
date: "April 23, 2017"
output: html_document
runtime: shiny
---
## **Introduction**

  This assignment is part of Developing Data Products course in Course era. 
  The Assignment is to create a shiny application. The following key features of Shiny 
  application are explored in this exercise.
  1. SelectInput box
  2. Adding a image to your application
  3. Using HTML tags like Table
  4. Using Plotly graphs and Ploty specific shiny render funtions
  5. TabPanels
  
### **Dataset Used**

  The original data set from <https://catalog.data.gov/dataset/crimes-2001-to-present-398a4> 
  reflects reported incidents of crime (with the exception of murders where data exists 
  for each victim) that occurred in the City of Chicago from 2001 to present, minus the most
  recent seven days. Data is extracted from the Chicago Police Department's CLEAR (Citizen Law
  Enforcement Analysis and Reporting) system. In order to protect the privacy of crime victims,
  addresses are shown at the block level only and specific locations are not identified.

### **User input**

    The application segments the crime data based on crime type and crime location and displays 
    the trend in chicago for the last 18 years. As a end user, you can visualize the trends across
    various types. If the user wishes to drill down on last 5 years, he/she can  select the time
    frame from the input drop-down and hit the **process** button. The plots are refreshed for 
    the selected time period. *Please not the change in x-axis as the result of the time frame selected*
    

```{r echo=FALSE}
    library(shiny)
    library(plotly)
    library(dplyr)
    library(RColorBrewer)
```

### **client code**

    The client has the image, the header, the selectinput drop-down, the **submit** button and **tab** controls.
    
```{r}    
    ui <- fluidPage(title = "Chicago City Crime",
            tags$table( tags$tr(tags$td(img(src= "CHICAGOPOLICE.jpg")),tags$td(h1("City of Chicago Crime Analysis"))),
                      tags$tr(tags$td( selectInput(inputId = "YearSel",
                                                   label = " Select the Time period for analysis",
                                                   choices = c("All years" = "0", "Last 5 years" ="1", "Last 10 years"="2", "Last 15 years"="3"),
                                                   multiple = F)))),
            submitButton( "Process"),
            
            tabsetPanel(              
                        tabPanel(title = "Top Ten Crime type",
                                 plotlyOutput("TopTenCrimesType")
                        ),
                        tabPanel(title = "Top Ten Crime locations",
                                 plotlyOutput("TopTenCrimesLocation")
                         )
             )
    )
    
```

### **Server side code**
    
    The server side, handles reading the data from the csv file, processing and rendering them into plots.
    
```{r}
    server <- function(input, output) {
       
        m <- list(
            l = 50,
            r = 50,
            b = 100,
            t = 100,
            pad = 4
        )
        
       Time_period <- reactive({
           input$YearSel
           }) 
       
        output$TopTenCrimesType <- renderPlotly({
           plot_data <- read.csv(".\\Top10Crimes.csv"
                                 ,stringsAsFactors = TRUE
                                 , header = TRUE)
           
           heading <- " City of Chicago Top 10 crimes by type in all years"
           if(Time_period() == "1") {
               plot_data <- filter(plot_data,Year >= max(Year) - 5 )
               heading <- " City of Chicago Top 10 crimes by type in last 5 years"
           }

           if(Time_period() == "2") {
               plot_data <- filter(plot_data,Year >= max(Year) - 10 )
               heading <- "City of Chicago Top 10 crimes by type in last 10 years"
           }

           if(Time_period() == "3"){
               plot_data <- filter(plot_data,Year >= max(Year) - 15 )
               heading <- " City of Chicago Top 10 crimes by type in last 15 years"
            }
           
       
                  # use plot to plot.
              plot_ly(data = plot_data,
                           x = ~Year,
                           y = ~count,
                           color = ~Primary.Type,
                           size = ~size,
                           colors = brewer.pal(10, "Paired"), 
                           type = 'scatter',
                           mode = 'markers',
                           width = 800,
                           height = 600,
                           marker = list( symbol = 'circle',
                                          sizemode = 'diameter',
                                          line = list( width = 2,
                                                       color = '#FFFFFF')),
                           text = ~paste('Year:', Year, '<br>Crime Type:', Primary.Type, '<br> Count:', count)
              )   %>%
                  layout( title =  heading 
                           , autosize = F
                            , margin = m
                            ,xaxis = list(title = 'Time Series in years',
                            range = c(min(plot_data$Year)-2, max(plot_data$Year)+2)
                          ),
                          yaxis = list(title = 'Crime count',
                                       range = c(min(plot_data$count), max(plot_data$count))
                          )
                  )
         
      })
       
      output$TopTenCrimesLocation <- renderPlotly({
        Crime_byLocation_data <-read.csv(".\\Crime_by_Location.csv"
                             ,stringsAsFactors = TRUE
                             , header = TRUE)  
        
        Crime_byLocation_data$size <- sqrt(Crime_byLocation_data$count * 2.666051223553066e-05 )
        
        m <- list(
            l = 50,
            r = 50,
            b = 100,
            t = 100,
            pad = 4
        )
        heading <- " City of Chicago Top 10 crimes by location in all years"
        if(Time_period() == "1") {
            Crime_byLocation_data <- filter(Crime_byLocation_data,Year >= max(Year) - 5 )
            heading <- " City of Chicago Top 10 crimes by location in last 5 years"
        }
        
        if(Time_period() == "2") {
            Crime_byLocation_data <- filter(Crime_byLocation_data,Year >= max(Year) - 10 )
            heading <- "City of Chicago Top 10 crimes by location in last 10 years"
        }
        
        if(Time_period() == "3"){
            Crime_byLocation_data <- filter(Crime_byLocation_data,Year >= max(Year) - 15 )
            heading <- " City of Chicago Top 10 crimes by location in last 15 years"
        }
        
            # use plot to plot.
        plot_ly(data = Crime_byLocation_data,
                x = ~Year,
                y = ~count,
                color = ~Location.Description,
                size = ~size,
                colors = brewer.pal(10, "Paired"), 
                type = 'scatter',
                mode = 'markers',
                width = 800,
                height = 600,
                marker = list( symbol = 'circle',
                               sizemode = 'diameter',
                               line = list( width = 2,
                                            color = '#FFFFFF')),
                text = ~paste('Year:', Year, '<br>Crime Type:', Location.Description, '<br> Count:', count)
        )   %>%
            layout( title = heading,autosize = F, margin = m, xaxis = list(title = 'Time Series in years',
                                range = c(min(Crime_byLocation_data$Year)-2, max(Crime_byLocation_data$Year)+2)
            ),
            yaxis = list(title = 'Crime count',
                         range = c(min(Crime_byLocation_data$count), max(Crime_byLocation_data$count))
            )
            )
      })
    }
    
    shinyApp(server = server, ui = ui)
```

### **Conclusion**

     The assignment is complete with input, output, reactive functionalities. Now the user has the ability to drill-down and visualize the crime data as needed.
